Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-07-07 attenuation.tex
finalizing results section

Based on this, we selected the GMPE proposed by \citet{Kalkan2004} for our analysis. This equation is of the form
% 
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\ 
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
% 
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, $V_S$ is the shear wave velocity of the site of interest, and the distance $r$ is given by
% 
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, .
\end{equation}
% 
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
% 
\begin{equation}
\begin{array}{lcrlcr} 
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,; 
	\nonumber
\end{array}
\end{equation}
% 
and $V_A$, $V_S$ and $h$ are
% 
\begin{equation}
\begin{array}{lcrl} 
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
% 
respectively. The value of $V_S$ is assumed to represent average surface rock sites.

>>>>

Based on this, we selected the GMPE proposed by \citet{Kalkan2004} for our analysis. This equation is of the form
% 
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\ 
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
% 
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, $V_S$ is the shear wave velocity of the site of interest, and the distance $r$ is given by
% 
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
% 
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
% 
\begin{equation}
\begin{array}{lcrlcr} 
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,; 
	\nonumber
\end{array}
\end{equation}
% 
and $V_A$, $V_S$ and $h$ are
% 
\begin{equation}
\begin{array}{lcrl} 
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
% 
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, in consistence with \citet{Zafarani2014}, we set the models in the hazard analysis to compute ground motions (i.e., PGAs) at distances no greater than 200 km.

### 2016-07-07 hazard.tex
finalizing results section

It is obvious from Eq.~\ref{eq:nnorm} that an important parameter in smoothing the seismicity is the correlation distance. Different values of $c$ have been used in past studies, ranging between 10 and 50 km. \citet{Frankel1995} and \citet{Boyd2008} assumed $c=50$~km, \citet{Barani2007} used $c=25$~km, and \citet{Foteva2006} used $c=10$ and 15 km. According to \citet{Mirzaei1997}, determining an appropriate value of $c$ depends on the crustal structure of the region, the distribution of stations, and the number and quality of records. We use an intermediate value of $c=30$~km. This selection is consistent with epicentral uncertainty for previous hazard zoning maps for Iran \citep[see][]{Zare2012}.

>>>>

It is obvious from Eq.~\ref{eq:nnorm} that an important parameter in smoothing the seismicity is the correlation distance. Different values of $c$ have been used in past studies, ranging between 10 and 50 km. \citet{Frankel1995} and \citet{Boyd2008} assumed a correlation distance of 50~km, \citet{Foteva2006} used 10 and 15 km, \citet{Barani2007} used 25~km, and \citet{Khodaverdian_2016_BSSA} used 40 km. According to \citet{Mirzaei1997}, determining an appropriate value of $c$ depends on the crustal structure of the region, the distribution of stations, and the number and quality of records. We use an intermediate value of $c=30$~km. This selection is consistent with epicentral uncertainty for previous hazard zoning maps for Iran \citep[see][]{Zare2012}.

### 2016-07-07 introduction.tex
finalizing results section

According to this tectonic setting and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968, Takin1972, Berberian1976}, and up to twenty to twenty-three seismotectonic provinces in the most elaborate ones \citep[e.g.,][]{Nowroozi1976, Tavakoli1999}. Contemporary, well-accepted models divide Iran into five major tectonic regions: Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran \citep[e.g.,][]{Mirzaei1998}. More recently, in a study aimed to develop a uniform earthquake catalog for Iran, \citet{Karimiparidari2013} further divided the Azerbaijan-Alborz region into two separate seismic zones: Azerbaijan and the Alborz Mountain Range, hereafter referred to simply as Alborz. We adopt the latter definition of six seismic seismic zones and focus our attention on those most relevant to the northern part of the country, Azerbaijan, Alborz and Kopeh-Dagh.

>>>>

According to this tectonic setting and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968, Takin1972, Berberian1976}, and up to twenty to twenty-three seismotectonic provinces in the most elaborate ones \citep[e.g.,][]{Nowroozi1976, Tavakoli1999}. Contemporary, well-accepted models divide Iran into five major tectonic regions: Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran \citep[e.g.,][]{Mirzaei1998}. More recently, in a study aimed to develop a uniform earthquake catalog for Iran, \citet{Karimiparidari2013} further divided the Azerbaijan-Alborz region into two separate seismic zones: Azerbaijan and the Alborz Mountain Range, hereafter referred to simply as Alborz. We adopt the latter definition of six seismic zones and focus our attention on those most relevant to the northern part of the country, Azerbaijan, Alborz and Kopeh-Dagh.

### 2016-07-07 parameters.tex
finalizing results section

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-07} 
    \caption{Annual probability of exceedance as a function of earthquake magnitude for the seismic zones in the region of interest. The three curves in red correspond to those seismic zones completely considered within the study area, whereas the ones in blue are the two additional contributing zones from the rest of the country. THe last curve in green corresponds to the complete region of interest, here referred to as the uniform model. The shaded range indicates the models' $\pm 1$ standard deviation.}
    \label{fig:annualp}
\end{figure*}

@@

The $a$-value is estimated based on the completeness year, the correlation distance, and the number of events per grid cell. We assume magnitude increments of $\Delta M = 0.1$, count events with magnitude $M_w > 3$, and smooth the number of events for each cell. This means each cell has a different $a$-value and thus the result is variable throughout our region of interest. On the other hand, we consider the $b$-value and $M_{\max}$ to be constant within each of the seismic zones in our study area. We compute $b$ and $M_{\max}$ using HA3 and the information from the catalog, which we divide in four sub-catalogs: 2 incomplete catalogs for the prehistorical (before 1000) and historical (between 1000 and 1799) event periods, and two complete catalogs for  early instrumental (between 1800 and 1899) and instrumental (between 1900 and 2015) event periods. In this process we use the values of $M_c$ shown in Table \ref{tab:completeness}.

We obtain $b$ and $M_{\max}$ for two different models. In the first model we obtain these parameters independently for each one of the five seismic zones influencing the hazard in northern Iran. That is, we obtain distinct values of $b$ and $M_{\max}$ for the main northern seismic zones of Azerbaijan, Alborz, and Kopeh Dagh, as well as for the portions of the Zagros and Central-East seismic zones contributing to our region of interest (see Fig.~\ref{fig:iran} for reference). In the second model we consider the whole study area as a uniform seismic zone. Admittedly, this assumption is unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. Computing results for this uniform model is, nonetheless, useful for the analysis of results and comparisons with other studies. Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained with HA3 for the five seismic zones and the uniform regional model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$; and Fig.~\ref{fig:annualp} shows the annual probability of exceedance as a function of earthquake magnitude.

>>>>

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-07} 
    \caption{Annual probability of exceedance as a function of earthquake magnitude for the seismic zones in the region of interest. The three curves in red correspond to those seismic zones completely considered within the study area, whereas the ones in blue are the two additional contributing zones from the rest of the country. The last curve in green corresponds to the complete region of interest, here referred to as the uniform model. The shaded range indicates the models' $\pm 1$ standard deviation.}
    \label{fig:annualp}
\end{figure*}

@@

The $a$-value is estimated based on the completeness year, the correlation distance, and the number of events per grid cell. We assume magnitude increments of $\Delta M = 0.1$, count events with magnitude $M_w > 3$, and smooth the number of events for each cell. This means each cell has a different $a$-value and thus the result is variable throughout our region of interest. On the other hand, we consider the $b$-value and $M_{\max}$ to be constant within each of the seismic zones in our study area. We compute $b$ and $M_{\max}$ using HA3 and the information from the catalog, which we divide in five sub-catalogs: one incomplete catalog for the prehistorical period before 1000, one incomplete catalog for the historical period between 1000 and 1799, one complete catalog for the early instrumental period between 1800 and 1899, and two complete catalogs for the instrumental periods between 1900 and 1999, and between 2000 and 2015. In this process we use the values of $M_c$ shown in Table \ref{tab:completeness}.

We obtain $b$ and $M_{\max}$ for two different models. In the first model we obtain these parameters independently for each one of the five seismic zones influencing the hazard in northern Iran. That is, we obtain distinct values of $b$ and $M_{\max}$ for the main northern seismic zones of Azerbaijan, Alborz, and Kopeh Dagh, as well as for the Zagros and Central-East seismic zones contributing to our region of interest (see Fig.~\ref{fig:iran} for reference). In the second model we consider the whole study area as a uniform seismic zone. Admittedly, this assumption is unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. Computing results for this uniform model is, nonetheless, useful for the analysis of results and comparisons with other studies. Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained with HA3 for the five seismic zones and the uniform regional model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$; and Fig.~\ref{fig:annualp} shows the annual probability of exceedance as a function of earthquake magnitude.

### 2016-07-07 results.tex
figures 12 and 13 added
finalizing results section

Fig.~\ref{fig:pga} shows the expected mean PGA values obtained using the selected attenuation relationship as a measured of gravity (g) throughout the region of interest for the different model combinations. As can be expected, the highest levels of ground motion are associated with the 2R5.5 and 2U5.5 models, and the lowest with the 10R6.5 and 10U6.5 models. In results for the R models shown Fig.~\ref{fig:pga}a, the highest levels of expected ground motion occur in the Kopeh Dagh region along the border with Turkmenistan, followed by ground motions in the north-central region west from Tehran and Karaj, and in the vicinity of the Piranshahr fault between the Zagros and Azerbaijan seismic regions (see Fig.~\ref{fig:selected}b for reference). The Azerbaijan seismic zone does not present equally high levels of expected ground motion. In the U models shown in Fig.~\ref{fig:pga}b, on the other hand, the highest levels of expected ground motion are observed precisely in the area corresponding to the Azerbaijan seismic region, especially in the vicinity of the North Tabriz Fault system (see Fig.~\ref{fig:selected} for reference).

@@

Overall, these results are consistent with the seismicity observed in Fig.~\ref{fig:catalog}. The R model results, however, seem to be more sensitive to the influence of large magnitude events observed in both the historical and modern seismicity maps, with concentration of strong earthquakes in the Alborz and Kopeh Dagh regions; whereas the U model results seem to be more sensitive to the influence of low-to-moderate magnitude events, especially in the Azerbaijan region. This may be the cause of

@@

We also note that the overall distribution of PGA values shown in Fig.~\ref{fig:pgaavgs}c are the ones that best agree with the recent study by \citet{Khodaverdian_2016_BSSA}. These estimates also agree, on average, with other previous studies. For the north-west seismic regions of Azerbaijan and Zagros, for instance, \citet{Tavakoli1999} predicts maximum mean PGA values of 0.45 g near the North Tabriz Fault and the North Tehran Fault zones, and 0.4 g around the city of Urmia, where we find equivalent values to be between 0.4 and 0.6 g, and about 0.35 g, respectively. The results from \citet{Vafaie2011} for the city of Tabriz, on the other hand, are slightly higher than the averages shown in Fig.~\ref{fig:pgaavgs}c and are closer to those we present in Fig.~\ref{fig:pga}b, which are in the order of 0.8 g and 0.6 for the models 2U5.5 and 10U5.5. In the north-central region of Alborz, \citet{Ghodrati2003} and \citet{Boostan2015} predict PGA values of the order 0.27--0.46 g and 0.42--0.48 for Tehran, where we found equivalent results to be between 0.3 and 0.45 g. Similarly, \citet{Abdi2013} reports expected PGA values in the vicinity of Karaj and Tehran to be on the order of 0.3 g. Results from \citet{Ghodrati2008} for the Gilan province around the city of Rasht are also of the same order of those we show in Figs.~\ref{fig:pga} and \ref{fig:pgaavgs}c. Finally, in for the Kopeh Dagh region, and in particular, for the city of Bojnurd in the nort-east province of Khorasan, \citet{Rahgozar2012} reports expected PGA values for return periods of 475 and 2475 years in range of 0.15--0.22 g and 0.28--0.49 g, respectively, where for magnitudes exceeding $M_w$ 6.5 we found equivalent values to be in the order of 0.2--0.4 and 0.4--0.6 g. Table \ref{tab:pga} shows a condensed version of these comparisons between our results and those of other previous studies for selected cities in northern Iran.

@@

Fig.~\ref{fig:hazardcurve} represents the hazard curve of different model for some of important northern cities in Iran.

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.4]{figures/pdf/HazardCurve.pdf} 
\caption{Seismic Hazard Curve of Northern Cities in Iran}
\label{fig:hazardcurve}
\end{figure*}

>>>>

Fig.~\ref{fig:pga} shows the expected mean PGA values obtained using the selected attenuation relationship as a measured of gravity (g) throughout the region of interest for the different model combinations. As can be expected, the highest levels of ground motion are associated with the 2R5.5 and 2U5.5 models, and the lowest with the 10R6.5 and 10U6.5 models. In results for the R models shown Fig.~\ref{fig:pga}a, the highest levels of expected ground motion occur in the Kopeh Dagh region along the border with Turkmenistan, and in a small area half-way between the city of Karaj and the Piranshahr fault, where a small concentration of events can be observed in Fig.~\ref{fig:catalog}b, as well as along the Piranshahr and Morvarid fault lines between the Zagros and Azerbaijan seismic regions (see Fig.~\ref{fig:selected}a for reference). The Azerbaijan seismic zone does not present equally high levels of expected ground motion. In the U models shown in Fig.~\ref{fig:pga}b, on the other hand, the highest levels of expected ground motion are observed precisely in the area corresponding to the Azerbaijan seismic region, especially in the vicinity of the North Tabriz Fault system (see Fig.~\ref{fig:selected} for reference).

@@

Overall, these results are consistent with the seismicity observed in Fig.~\ref{fig:catalog}. The R model results, however, seem to be more sensitive to the influence of large magnitude events observed in both the historical and modern seismicity maps, with concentration of strong earthquakes in the Alborz and Kopeh Dagh regions; whereas the U model results seem to be more sensitive to the influence of low-to-moderate magnitude events, especially in the Azerbaijan region. This is because of the higher $b$-value obtained for the Azerbaijan region in the R model as opposed to the lower value for the U model.

@@

We also note that the overall distribution of PGA values shown in Fig.~\ref{fig:pgaavgs}c are the ones that best agree with the recent study by \citet{Khodaverdian_2016_BSSA}. These estimates also agree, on average, with other previous studies. For the north-west seismic regions of Azerbaijan and Zagros, for instance, \citet{Tavakoli1999} predicts maximum mean PGA values of 0.45 g near the North Tabriz Fault and the North Tehran Fault zones, and 0.4 g around the city of Urmia, where we find equivalent values to be between 0.4 and 0.6 g, and about 0.35 g, respectively. The results from \citet{Vafaie2011} for the city of Tabriz, on the other hand, are slightly higher than the averages shown in Fig.~\ref{fig:pgaavgs}c and are closer to those we present in Fig.~\ref{fig:pga}b, which are in the order of 0.8 g and 0.6 for the models 2U5.5 and 10U5.5. In the north-central region of Alborz, \citet{Ghodrati2003} and \citet{Boostan2015} predict PGA values of the order 0.27--0.46 g and 0.42--0.48 for Tehran, where we found equivalent results to be between 0.3 and 0.45 g. Similarly, \citet{Abdi2013} reports expected PGA values in the vicinity of Karaj and Tehran to be on the order of 0.3 g. Results from \citet{Ghodrati2008} for the Gilan province around the city of Rasht are also of the same order of those we show in Figs.~\ref{fig:pga} and \ref{fig:pgaavgs}c. Finally, for the Kopeh Dagh region, and in particular, for the city of Bojnurd in the northeast province of Khorasan, \citet{Rahgozar2012} reports expected PGA values for return periods of 475 and 2475 years in range of 0.15--0.22 g and 0.28--0.49 g, respectively, where for magnitudes exceeding $M_w$ 6.5 we found equivalent values to be in the order of 0.2--0.4 and 0.4--0.6 g.

\begin{figure*}[th!]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-12} 
    \caption{Hazard curves representing the annual probability of exceedance as function of peak ground acceleration for six cities in northern Iran based on the four basic models considered in this study, namely R5.5, R6.5, U5.5 and U6.5.}
    \label{fig:hazardcurve}
\end{figure*}

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-13} 
    \caption{Hazard curve comparison between the models considered in this study and the results of two previous studies by \citet{Ghodrati2003} and \citet{Boostan2015} for the city of Tehran. Here, the two previous studies just mentioned are identified with the codes Gh03 and Bo15, respectively.}
    \label{fig:tehran}
\end{figure}

Table \ref{tab:pga} shows a condensed version of these comparisons between our results and those of other previous studies for selected cities in northern Iran. In addition, we also looked into the more complete picture of the annual probabilities of exceedance as functions of PGA for six cities (all the same as in Table \ref{tab:pga} except Rasht). This is shown in Fig.~\ref{fig:hazardcurve} for the four basic models considered before, namely R5.5, R6.5, U5.5 and U6.5, each also including upper and lower ranges of one standard deviation. These hazard curves exemplify again the differences between the R and U models, which are more prominent in the northwest cities of Tabriz and Ardabil, where the U models yield higher expected values of PGA for any given annual probability of exceedance. Minor differences are also present in the central cities of Karaj and Tehran, whereas the cities of Urmia and Mashhad in the Zagros and Kopeh Dagh seismic zones present similar hazard levels between the two models (R and U). This is consistent with our previous observations regarding Figs.~\ref{fig:pga}, \ref{fig:pgadiff} and \ref{fig:pgaavgs}.

Last, we pay additional attention to the capital city of Tehran and compare hazard...

@@
